Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-13 paper.qmd
Updated Figure 12

  ggplot(data = digit_df, aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = ncol) +
  ggtitle(title_text) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = palette) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = hjust,
                                  size = 12))

@@

error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
                           aes(x = a1, 
                               y = MSE, 
                               colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a, b, e that show two close clusters are universally suboptimal. Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout e has small separation with oddly shaped clusters. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

>>>>

  ggplot(data = digit_df, aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = ncol) +
  ggtitle(title_text) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = palette) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = hjust,
                                  size = 15))

@@

error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
                           aes(x = a1, 
                               y = MSE, 
                               colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 12))

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a, b, e that show two close clusters are universally suboptimal. Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout e has small separation with oddly shaped clusters. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 100%
